Three-Dimensional Confocal Raman Temperature Characterization of Electrokinetically Pumped Microchannels

### 2018-12-10 paper.tex
ultima version

\title{Three-Dimensional Temperature Characterization of Electrokinetically Pumped Microchannels}
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.

@@

\begin{abstract}
A novel method for non invasive, three dimensional temperature characterization in microfluidic devices is presented. A specially designed confocal microscope was constructed and used to measure water temperature by sensing the Raman spectrum variations of the liquid. A fiber coupled laser beam is used to pump the sample with 25~mW of optical power at 405~nm. This together with the confocal character of the device allows a 0.8~K temperature resolution and a 9~$\mu$m axial resolution using a 1 second integration time and a high numeric aperture microscope objective. This features make temperature profiling in all dimensions possible, in contraposition with previous methods where the information present in the height of the channel is lost. Using this technique different geometries of PDMS microchannels sealed with a 150~$\mu$m glass coverslip where studied, showing that heat flow trough the glass is the dominating dissipation mechanism and defines the maximum temperature in the channel. The results show good agreement with previous work found in literature.
\keywords{Raman thermometer \and electroosmotic flow \and spatial resolution}
\end{abstract}

@@

Typical characteristic dimensions in microfluidic devices are in the range $\sim10-100\, \mathrm{\mu m}$. In these devices it is generally necessary to either move the liquid or move species in solution from one point to another in the microfluidic network. To do so, one needs a pumping or species transport strategy. In many cases this is done via electroosmotic pumping or electrophoresis \cite{hunter2001,lyklema1995}. In any of these cases, an electric current is passed through the liquid, the applied power has to be dissipated. Depending on the device dissipation efficiency, the liquid temperature will rise more or less. In any case, as these devices are often used in biological applications \cite{tian2008}, temperature is a critical parameter that has to be monitored and controlled. Moreover, even in non biological applications, the electroosmotic and electrophoretic mobility are a function of temperature \cite{tang2006} so that temperature rise may lead to unwanted pumping or electrophoretic behaviour. The readers are referred to the reviews by Xuan~\cite{xuan2008} and Cetin and Li~\cite{cetin2008} for a deeper description of joule heating in electrokinetic flows. 
Different techniques were developed to measure temperature inside microchannels. A common approach is to introduce in the flow a dye with a temperature dependent property~\cite{tang2006,ross2001,erickson2003,dye1,dye2}. The measurement of this property with the proper calibration is used as a thermometer. This is an invasive method, with the disadvantage that the dye can interfere with the chemical reactions occurring in the channel or vice versa. This makes the method not useful for online monitoring. Moreover, the measurement ``integrates'' the temperature in the channel height, not allowing for the measurement of temperature profiles normal to the channel bottom.

Walrafen et al \cite{walrafen1} discovered that the infrared spectrum of water shows temperature dependent qualities and measured the existence of an isosbestic point around 3425~$\mathrm{cm^{-1}}$. This molecular property can be sensed using Raman Scattering to get a temperature dependent signal that is intrinsic of water. This method was proposed and proved useful in microchannels \cite{raman1}\cite{raman2}\cite{raman3} where an epi-illumination microscope was adapted to measure Raman Scattering by exciting the water with a visible laser. As water absorption is low in the visible spectrum \cite{absorption} this method proved to be non-invasive. In this work we propose an improvement over these methods were spatial resolution is enhanced by using a confocal collection scheme. This is achieved by focusing the collected Raman signal in a optical fibre (instead of the traditional pinhole used in most confocal microscopes). This result in a device capable of three dimensional maping of the channel, allowing the study of thermal gradients in every direction. Also, the use of a 405~nm excitation beam maximizes temperature signal as Raman emission increases strongly with frequency \cite{faris}, besides being a cheap wavelength. Finally, a dark field illumination scheme is used to get scattering or fluorescence images of the sample at the same time temperature measurements take place. This allowed to perform temperature maps that characterize the heat flow in the channel and the geometric and construction parameters that dominate heat dissipation. Our results show good agreement with those found in the literature.

@@

At the other end of the multimode optical fiber the Raman signal beam is divided in the two spectral bands necessary for the temperature measurement and then detected. For this purpose a transmission diffraction grating (DG) with 300~l/mm and a 200~mm achromatic lens (SL) is used as a simple spectrometer. A D shaped mirror (DM) is used in the lens focal plane to separate the two spectral bands. The latter is mounted in a micrometric translation stage which allows a fine tunning of the separation wavelength which must be matched to the isosbectic point in order to maximize the temperature sensitivity of the method. After this both reflected and transmited beam are directed to the detectors. A 75 mm lens (LA and LB) is used in each path to make a one on one image of DM in each detector. A longpass filter centered at 450~nm (FA and FB) is used in each channel to reject the laser wavelength. Silver mirrors M1 to M6 are used for beam steering and aligning. The detectors are two photon counting modules (Hamamatsu H7828) providing a quantum efficiency of around 15\% at the desired wavelength. The devices are designed to produce an amplified and shaped square pulse for each detected photon simplifying the photon counting procedure. The pulses are then counted in a selectable time window using an on board DAQ device on a PC. Counting rates of around $1.5 \cdot 10^5 \,\,\mathrm{s}^{-1}$ are achieved at each channel. After this, the final signal is computed at software level as
\begin{equation}\label{eq:S}
S = \frac{B – A}{A + B}
\end{equation}
where A and B are the counts of each channel. The temperature of the sample is obtained from this calculation after proper calibration.  It is important to notice that $S$ is computed as a ratio, making the calibration independent of experimental parameters such as the excitation power or the integration time.

Additionally the diffraction grating (DG) is mounted in a motorized rotation stage.  This is not completely necessary to the Raman measurements but allows using the device as a spectrometer by measuring the signal as a function of the incidence angle on the grating and then taking the numeric derivative of the data. This simplifies the procedure of choosing the isosbectic point and allows checking the detected spectrum to match the one of water, confirming the proper alignment of the microscope.

@@

To characterize the technique a series of Raman spectra were measured at different temperatures. This was performed by using a pure water sample in the temperature controlled cell and by scanning the angle on the diffraction grating. After taking the numeric derivative, the recovered spectra is shown in figure \ref{fig:spectra}. Results show an isosbestic point around 470 nm in agreement with previous measurements\cite{walrafen1}. As temperature rise, the shorter wavelength components of the Raman spectrum reduce their intensity while the longer wavelength part slightly increases. The total spectrum collected in each detector (channel A and B) is shaded in light blue and red. This measurements are used to set the angle at which the spectrum is split exactly at the isosbectic point, leading to the maximum temperature sensibility.

@@

Having set the spectrometer properly, a calibration is made by computing the quantity $S$, as defined in equation \ref{eq:S}, as a function of temperature. The result from this experiment is shown in figure \ref{fig:calib} where a linear behavior is exhibited in the range of interest. A linear fit is finally used to obtain the calibration curve. With this calibration and by measuring the noise in the temperature signal  the resolution of the method can be estimated. The result of this experiment is consistent with the spread of the data around the linear fit in figure \ref{fig:calib} with a value of 0.8~K for a integration time of one second.

@@

With this results a full characterization of the thermal properties of the microchannels under electroosmotic flow is possible. First of all, the the temperature rise at \textit{turn on} is studied, from ambient temperature (when the electric field is turned off) to the equilibrium temperature once the flow is established. The result of this experiment for a channel with dimensions b = 80~$\mathrm{\mu m}$, h = 100~$\mathrm{\mu m}$ and $L$ = 20~mm is shown if figure \ref{fig:temporal}. Electric field is turned on at $t = 0$. The flow velocity was measured to be 1~mm/s. A temperature rise of 12~K is observer in a short time (around 10 seconds) for two different positions in the channel 5~mm apart. This timing is independent of the position of the channel, showing no mass transport effects.

@@

Acá la cuentita que habia hecho yo quedó obsoleta dado el modelo de Martín. 
%A simple calculation considering that all heat flows trough the floor of the channel (glass) can be useful to estimate the required temperature gradient needed to reach a stationary regime where the extracted heat compensates Joule effect. For a typical dissipated power of 40 mW and a channel of dimensions $b = h = 100$ $\mathrm{\mu m}$ and $L$ = 20~mm the required temperature difference between the ceiling and the floor is 
%
%\begin{equation}
%\Delta T = \frac{Ph}{kLb} = 1.7\,\, \mathrm{K},
%\end{equation}

@@

To confirm this hypothesis curves of temperature rise as function of dissipated power per unit glass area were measured for different channel geometries. The result of this experiment is plotted in figure \ref{fig:rectas}. A linear fit was performed for each data set. Results show that all different geometries have a similar slope, proving that glass area is the most relevant parameter defining heat dissipation out of our channels. This is also consistent with theoretical and experimental results from other groups \cite{} where temperature rise between all PDMS and PDMS/glass channels was compared, obtaining a much lower heating effect in the latter.

@@

A novel technique was presented that proven useful to characterize Joule heating effects in eletrokinetically pumped microchannels with three-dimensional resolution. The thermometer, that works using the temperature dependence in the Raman spectrum of liquid water, improves from previous methods by using a confocal collection scheme. This allows a much higher spatial resolution, allowing thermal characterization in the direction normal to the channels base, which is impossible in other methods where the complete channel height is integrated.

 A temperature resolution of about 0.8 K was achieved in a one second integration time using 25~mW of optical power at the sample provided from a 405~nm fiber coupled laser beam. Temperature profiles were measured in different channel geometries with results compatible with a simple heat flow model. From this experiments we conclude that heat dissipation is occurring through the glass floor of the channel in accordance with theoretical and experimental results found in literature. As temperature rise is an important parameter to control in many of the applications of these devices, the method provides a complete three dimensional characterization of Joule heating effects that allows for improved channel design and temperature monitoring.

>>>>

\title{Three-Dimensional Confocal Raman Temperature Characterization of Electrokinetically Pumped Microchannels}
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.

@@

\begin{abstract}
A novel method for non invasive, three dimensional temperature characterization in microfluidic devices is presented. A specially designed confocal microscope was constructed and used to measure water temperature by sensing the Raman spectrum variations of the liquid. This is achieved by splitting the spectrum in the isosbestic point and detecting with two photomultiplier tubes. The difference in the signals of each detector divided by their sum shows a linear dependence with temperature. A fiber coupled laser beam is used to pump the sample with 25~mW of optical power at 405~nm. This together with the confocal character of the device allows a 0.8~K temperature resolution and a 9~$\mu$m axial resolution using a 1 second integration time and a high numeric aperture microscope objective. This features make temperature profiling in all dimensions possible, in contraposition with previous methods where the information present in the height of the channel is lost and the whole spectrum needed to be recovered before computing the sample temperature. Using this technique different geometries of PDMS microchannels sealed with a 150~$\mu$m glass coverslip where studied, showing that heat flow trough the glass is the dominating dissipation mechanism and defines the maximum temperature in the channel. The results show good agreement with previous work found in literature.
\keywords{Raman thermometer \and electroosmotic flow \and spatial resolution}
\end{abstract}

@@

Typical characteristic dimensions in microfluidic devices are in the range $\sim10-100\, \mathrm{\mu m}$. In these devices it is generally necessary to either move the liquid or move species in solution from one point to another in the microfluidic network. To do so, one needs a pumping or species transport strategy. In many cases this is done via electroosmotic pumping or electrophoresis \cite{hunter2001,lyklema1995}. In any of the latter, an electric current is passed through the liquid and the applied power has to be dissipated. Depending on the device dissipation efficiency, the liquid temperature will rise more or less. In any case, as these devices are often used in biological applications \cite{tian2008}, temperature is a critical parameter that has to be monitored and controlled. Moreover, even in non biological applications, the electroosmotic and electrophoretic mobility are a function of temperature \cite{tang2006} so that temperature rise may lead to unwanted pumping or electrophoretic behaviour. The readers are referred to the reviews by Xuan~\cite{xuan2008} and Cetin and Li~\cite{cetin2008} for a deeper description of joule heating in electrokinetic flows. 
Different techniques were developed to measure temperature inside microchannels. A common approach is to introduce in the flow a dye with a temperature dependent property~\cite{tang2006,ross2001,erickson2003,dye1,dye2}. The measurement of this property with the proper calibration is used as a thermometer. This is an invasive method, with the disadvantage that the dye can interfere with the chemical reactions occurring in the channel or vice versa. This makes the method not useful for online monitoring. Moreover, the measurement ``integrates'' the temperature in the channel height, not allowing for the measurement of temperature profiles normal to the channel bottom.

Walrafen et al \cite{walrafen1} discovered that the infrared spectrum of water shows temperature dependent qualities and measured the existence of an isosbestic point around 3425~$\mathrm{cm^{-1}}$. This molecular property can be sensed using Raman Scattering to get a temperature dependent signal that is intrinsic of water. This method was proposed and proved useful in microchannels \cite{raman1,raman2,raman3} where an epi-illumination microscope was adapted to measure the Raman spectrum by exciting the water with a visible laser. Temperature is then recovered by analizing these spectra. As water absorption is low in visible wavelenghts \cite{absorption} this method proved to be non-invasive. In this work we propose an improvement over these methods were spatial resolution is enhanced by using a confocal collection scheme and measurement velocity is increased by splitting the spectrum in two detectors. This results in a device capable of three dimensional maping of the channel, allowing the study of thermal gradients in every direction which is also faster as it eliminates the need of measuring the whole spectrum at each position. Additionally, the use of a 405~nm excitation beam maximizes temperature signal as Raman emission increases strongly with frequency \cite{faris}, besides being a cheap wavelength. Finally, a dark field illumination scheme is used to get scattering or fluorescence images of the sample at the same time temperature measurements take place. This allowed to perform temperature maps that characterize the heat flow in the channel and the geometric and construction parameters that dominate heat dissipation. Our results show good agreement with those found in the literature.

@@

At the other end of the multimode optical fiber the Raman signal beam is divided in the two spectral bands necessary for the temperature measurement and then detected. For this purpose a transmission diffraction grating (DG) with 300~l/mm and a 200~mm achromatic lens (SL) is used as a simple spectrometer. A D shaped mirror (DM) is used in the lens focal plane to separate the two spectral bands. The latter is mounted in a micrometric translation stage which allows a fine tunning of the separation wavelength which must be matched to the isosbestic point in order to maximize the temperature sensitivity of the method. After this both reflected and transmited beam are directed to the detectors. A 75 mm lens (LA and LB) is used in each path to make a one on one image of DM in each detector. A longpass filter centered at 450~nm (FA and FB) is used in each channel to reject the laser wavelength. Silver mirrors M1 to M6 are used for beam steering and aligning. The detectors are two photon counting modules (Hamamatsu H7828) providing a quantum efficiency of around 15\% at the desired wavelength. The devices are designed to produce an amplified and shaped square pulse for each detected photon simplifying the photon counting procedure. The pulses are then counted in a selectable time window using an on board DAQ device on a PC. Counting rates of around $1.5 \cdot 10^5 \,\,\mathrm{s}^{-1}$ are achieved at each channel. After this, the final signal is computed at software level as
\begin{equation}\label{eq:S}
S = \frac{B – A}{A + B}
\end{equation}
where A and B are the counts of each channel. The temperature of the sample is obtained from this calculation after proper calibration.  It is important to notice that $S$ is computed as a ratio, making the calibration independent of experimental parameters such as the excitation power or the integration time.

Additionally the diffraction grating (DG) is mounted in a motorized rotation stage.  This is not completely necessary to the Raman measurements but allows using the device as a spectrometer by measuring the signal as a function of the incidence angle on the grating and then taking the numeric derivative of the data. This simplifies the procedure of choosing the isosbestic point and allows checking the detected spectrum to match the one of water, confirming the proper alignment of the microscope.

@@

To characterize the technique a series of Raman spectra were measured at different temperatures. This was performed by using a pure water sample in the temperature controlled cell and by scanning the angle on the diffraction grating. After taking the numeric derivative, the recovered spectra are shown in figure \ref{fig:spectra}. Results show an isosbestic point around 470 nm in agreement with previous measurements\cite{walrafen1}. As temperature rise, the shorter wavelength components of the Raman spectrum reduce their intensity while the longer wavelength part slightly increases. The total spectrum collected in each detector (channel A and B) is shaded in light blue and red. This measurements are used to set the angle at which the spectrum is split exactly at the isosbestic point, leading to the maximum temperature sensibility.

@@

Having set the spectrometer properly, a calibration is made by computing the quantity $S$, as defined in equation \ref{eq:S}, as a function of temperature. The result from this experiment is shown in figure \ref{fig:calib} where a linear behavior is exhibited in the range of interest. A linear fit is finally used to obtain the calibration curve. With this calibration, and by measuring the noise in the temperature signal, the resolution of the method can be estimated. The result of this experiment is consistent with the spread of the data around the linear fit in figure \ref{fig:calib} with a value of 0.8~K for a integration time of one second.

@@

With this results a full characterization of the thermal properties of the microchannels under electroosmotic flow is possible. First of all, the the temperature rise at \textit{turn on} is studied, from ambient temperature (when the electric field is turned off) to the equilibrium temperature once the flow is established. The result of this experiment for a channel with dimensions b = 80~$\mathrm{\mu m}$, h = 100~$\mathrm{\mu m}$ and $L$ = 20~mm is shown if figure \ref{fig:temporal}. Electric field is turned on at $t = 0$. The flow velocity was measured to be 1~mm/s. A temperature rise of 12~K is observer in a short time (around 10 seconds) for two different positions in the channel 5~mm apart. This timing is independent of the position of the channel, showing no appreciable mass transport effects.

@@

%Acá la cuentita que habia hecho yo quedó obsoleta dado el modelo de Martín. 
%A simple calculation considering that all heat flows trough the floor of the channel (glass) can be useful to estimate the required temperature gradient needed to reach a stationary regime where the extracted heat compensates Joule effect. For a typical dissipated power of 40 mW and a channel of dimensions $b = h = 100$ $\mathrm{\mu m}$ and $L$ = 20~mm the required temperature difference between the ceiling and the floor is 
%
%\begin{equation}
%\Delta T = \frac{Ph}{kLb} = 1.7\,\, \mathrm{K},
%\end{equation}

@@

Finally, curves of temperature rise as function of dissipated power per unit glass area were measured for different channel geometries. The result of this experiment is plotted in figure \ref{fig:rectas}. A linear fit was performed for each data set. Results show that all different geometries have a similar slope, proving that glass area is the most relevant parameter defining heat dissipation out of our channels. This is also consistent with theoretical and experimental results from other groups \cite{erickson2003} where temperature rise between all PDMS and PDMS/glass channels was compared, obtaining a much lower heating effect in the latter.

@@

A novel technique was presented that proven useful to characterize Joule heating effects in eletrokinetically pumped microchannels with three-dimensional resolution. The thermometer, that works using the temperature dependence in the Raman spectrum of liquid water, improves from previous methods by using a confocal collection scheme and by splitting the spectrum at the isosbestic point, measuring the resulting spectral bands in two separate detectors. The resulting signal, which is linearly dependent with temperature, allows to characterize microfluidic devices with a high spatial resolution in a fast manner.

 A temperature resolution of about 0.8 K was achieved in a one second integration time using 25~mW of optical power at the sample provided from a 405~nm fiber coupled laser beam. Temperature profiles were measured in different channel geometries with results compatible with a simple heat flow model. From these experiments we conclude that heat dissipation is occurring through the glass floor of the channel in accordance with theoretical and experimental results found in literature. As temperature rise is an important parameter to control in many of the applications of these devices, the method presented here provides a complete three dimensional characterization of Joule heating effects that allows for improved channel design and temperature monitoring.
